---
title: "Honors Thesis"
author: "Dhanesha Mirpuri"
date: "May 2022"
output: html_notebook
---

```{r}
library("tidyverse")
library("AER")
library("lmtest")
library("sandwich")
library("ggplot2")
library("stargazer")
```

```{r}
data = read.table("/Users/dhaneshamirpuri/Documents/maindata6.csv", header = TRUE, fill = TRUE, sep = ',')
```

cleaning the dataset: 
- changing everything necessary to numerics
- replacing N/A's with 0's 

```{r}
data$etrincometaxespretaxincome[is.na(data$etrincometaxespretaxincome)] = 0
data$etrincometaxespretaxincome[data$etrincometaxespretaxincome == "#DIV/0!"] = 0
data$etrincometaxespretaxincome[is.na(data$etrincometaxespretaxincome)] = 0
data$etrincometaxespretaxincome = as.numeric(data$etrincometaxespretaxincome)
data$spending_on_tax = as.numeric(data$spending_on_tax)
data$totalassets = as.numeric(data$totalassets)
data$numberofemployees = as.numeric(data$numberofemployees)
data$totaldebt = as.numeric(data$totaldebt)
data$totalinventory = as.numeric(data$totalinventory)
data$pretaxincome = as.numeric(data$pretaxincome)
data$incometaxes = as.numeric(data$incometaxes)
data$propertyplantandequitment = as.numeric(data$propertyplantandequitment)
data$rateofassets = as.numeric(data$rateofassets)
data$researchanddevelopmentexpenses = as.numeric(data$researchanddevelopmentexpenses)
data$leverageratiooftotaldebttototala = as.numeric(data$leverageratiooftotaldebttototala)
data$capitalintensityratioofpropertyp = as.numeric(data$capitalintensityratioofpropertyp)
data$inventoryintensityratioofinvento = as.numeric(data$inventoryintensityratioofinvento)
data$totspending = as.numeric(data$totspending)

data

```
changing names of variables to make it easier

#why didnt this work - i have to repeat for every data set??
```{r}
etr = data$etrincometaxespretaxincome
taxlob = data$spending_on_tax
lob = data$totspending
assets = data$totalassets
emp = data$numberofemployees
debt = data$totaldebt
invt = data$totalassets
pretax = data$pretaxincome
inctax = data$incometaxes
ppe = data$propertyplantandequitment
roa = data$rateofassets
rd = data$researchanddevelopmentexpenses
levrat = data$leverageratiooftotaldebttototala
caprat = data$capitalintensityratioofpropertyp
invrat = data$inventoryintensityratioofinvento
```


after using histograms to check for outliers, dropping values that are less than -1 and greater than 1 for etr

```{r}
data2 = filter(data, etr <= 1 & etr >= -1)
data2
```

graphing etr post dropping outliers to confirm normal distribution

```{r}
etr2 = data2$etr*100
etrhist = hist(etr2, main = "Effective Tax Rate", xlab = "Effective Tax Rate (%)", breaks = 20, na.rm = TRUE)
```

testing if its skewed distribution
```{r}
median(data2$totspending, na.rm = TRUE)
mean(data2$totspending, na.rm = TRUE)
sqrt(var(data2$totspending, na.rm = TRUE))
```

Histograms to see distributions 

```{r}
dT1 = filter(data2, totspending <= 501500)
dT2 = filter(data2, totspending > 501500)
totlobbying1 = hist(dT1$totspending, main = "Lobbying Expenditure Distribution", xlab = "Amount Lobbied", breaks = 20, na.rm = TRUE)
totlobbying2 = hist(dT2$totspending, main = "Lobbying Expenditure Distribution", xlab = "Amount Lobbied", breaks = 20, na.rm = TRUE)
```

```{r}
dT1 = filter(data2, totspending >=1 & totspending <= 100000)
dT2 = filter(data2, totspending >= 10001 & totspending <= 30000)
dT3 = filter(data2, totspending >= 30001 & totspending <= 50000)
dT4 = filter(data2, totspending >= 50001 & totspending <= 70000)
dT5 = filter(data2, totspending >= 70001 & totspending <= 1000000)
dT6 = filter(data2, totspending >= 1000001 & totspending <= 25000000)
dT7 = filter(data2, totspending >= 25000001 & totspending <= 300000000)
dT8 = filter(data2, totspending >= 300000000)
dT_no_1 = filter(data2, totspending >=1 & totspending <= 2500000)

d_with_zero = filter(data2, totspending <2500000)
d_without_zero = filter(data2, totspending >=1 & totspending <= 2500000)
```

Histograms 
```{r}
totlobbying = hist(d_with_zero$totspending, main = "Lobbying Expenditure Distribution", xlab = "Amount Lobbied", frequency = TRUE, breaks = 20, na.rm = TRUE)
totlobbying2 = hist(d_without_zero$totspending, main = "Lobbying Expenditure Distribution", xlab = "Amount Lobbied", frequency = TRUE, breaks = 20, na.rm = TRUE)
```
```{r}
d_with_zero$year = as.numeric(d_with_zero$year)
d_without_zero$year = as.numeric(d_without_zero$year)
```

Filtering data by years (entire dataset)

```{r}
d2000 = filter(d_with_zero, year == 2000)
d2001 = filter(d_with_zero, year == 2001)
d2002 = filter(d_with_zero, year == 2002)
d2003 = filter(d_with_zero, year == 2003)
d2004 = filter(d_with_zero, year == 2004)
d2005 = filter(d_with_zero, year == 2005)
d2006 = filter(d_with_zero, year == 2006)
d2007 = filter(d_with_zero, year == 2007)
d2008 = filter(d_with_zero, year == 2008)
d2009 = filter(d_with_zero, year == 2009)
d2010 = filter(d_with_zero, year == 2010)
d2011 = filter(d_with_zero, year == 2011)
d2012 = filter(d_with_zero, year == 2012)
d2013 = filter(d_with_zero, year == 2013)
d2014 = filter(d_with_zero, year == 2014)
d2015 = filter(d_with_zero, year == 2015)
d2016 = filter(d_with_zero, year == 2016)
d2017 = filter(d_with_zero, year == 2017)
d2018 = filter(d_with_zero, year == 2018)
d2019 = filter(d_with_zero, year == 2019)

avglist = list(
  mean(d2000$totspending, na.rm= TRUE),
  mean(d2001$totspending, na.rm= TRUE),
  mean(d2002$totspending, na.rm= TRUE),
  mean(d2003$totspending, na.rm= TRUE),
  mean(d2004$totspending, na.rm= TRUE),
  mean(d2005$totspending, na.rm= TRUE),
  mean(d2006$totspending, na.rm= TRUE),
  mean(d2007$totspending, na.rm= TRUE),
  mean(d2008$totspending, na.rm= TRUE),
  mean(d2008$totspending, na.rm= TRUE),
  mean(d2010$totspending, na.rm= TRUE),
  mean(d2011$totspending, na.rm= TRUE),
  mean(d2012$totspending, na.rm= TRUE),
  mean(d2013$totspending, na.rm= TRUE),
  mean(d2014$totspending, na.rm= TRUE),
  mean(d2015$totspending, na.rm= TRUE),
  mean(d2016$totspending, na.rm= TRUE),
  mean(d2017$totspending, na.rm= TRUE),
  mean(d2018$totspending, na.rm= TRUE),
  mean(d2019$totspending, na.rm= TRUE))

avglist = as.numeric(avglist)
avglist
```

Descriptive statistics
```{r}
stargazer(d, type = "text", title="Descriptive statistics", digits=1, out="table1.txt")
stargazer(d_with_zero, type = "text", title="Descriptive statistics", digits=1, out="table1.txt")
stargazer(d_without_zero, type = "text", title="Descriptive statistics", digits=1, out="table1.txt")
```



Running individual regressions on etr and each variable to get some idea of correlation/relationship
(Kind of tells me which variables I need to log transform from the original dataset)

```{r}
etrreg1 = lm(etr ~ lob, d)
summary(etrreg1)

etrreg2 = lm(etr ~ taxlob, d)
summary(etrreg2)

etrreg3 = lm(etr ~ assets, d)
summary(etrreg3)

etrreg4 = lm(etr ~ emp, d)
summary(etrreg4)

etrreg5 = lm(etr ~ roa, d)
summary(etrreg5)

etrreg6 = lm(etr ~ rd, d)
summary(etrreg6)

etrreg7 = lm(etr ~ levrat, d)
summary(etrreg7)

etrreg8 = lm(etr ~ caprat, d)
summary(etrreg8)

etrreg9 = lm(etr ~ invrat, d)
summary(etrreg9)

etrreg10 = lm(etr ~log1p(lob),d )


```
Normal Regressions with total lobbying (both log and not logged) with both datasets
Deciding to log the positive data/lobbying variables

 log1p()

```{r}
tot1 = lm(etr ~ lob, d_with_zero)
summary(tot1)
coeftest(tot1,df=Inf,vcov=vcovHC(tot1,type = "HC1"))

tot2 = lm(etr ~ lob, d_without_zero)
summary(tot2)
coeftest(tot2,df=Inf,vcov=vcovHC(tot2,type = "HC1"))

tot3 = lm(etr ~ log1p(lob), d)
summary(tot3)
coeftest(tot3,df=Inf,vcov=vcovHC(tot3,type = "HC1"))
```

```{r}
tot4 = lm(etr ~ lob + assets , d_with_zero)
summary(tot4)
coeftest(tot4,df=Inf,vcov=vcovHC(tot4,type = "HC1"))

tot5 = lm(etr ~ lob + assets , d_without_zero)
summary(tot5)
coeftest(tot5,df=Inf,vcov=vcovHC(tot5,type = "HC1"))

tot6 = lm(etr ~ log1p(lob) + assets, d)
summary(tot6)
coeftest(tot6,df=Inf,vcov=vcovHC(tot6,type = "HC1"))
```
```{r}
tot7 = lm(etr ~ lob + assets + emp , d_with_zero)
summary(tot7)
coeftest(tot7,df=Inf,vcov=vcovHC(tot7,type = "HC1"))

tot8 = lm(etr ~ lob + assets +emp, d_without_zero)
summary(tot8)
coeftest(tot8,df=Inf,vcov=vcovHC(tot8,type = "HC1"))

tot9 = lm(etr ~ log1p(lob) + assets + log1p(emp), d)
summary(tot9)
coeftest(tot9,df=Inf,vcov=vcovHC(tot9,type = "HC1"))
```

```{r}
tot10 = lm(etr ~ lob + assets + emp + roa , d_with_zero)
summary(tot10)
coeftest(tot10,df=Inf,vcov=vcovHC(tot10,type = "HC1"))

tot11 = lm(etr ~ lob + assets + emp + roa, d_without_zero)
summary(tot11)
coeftest(tot11,df=Inf,vcov=vcovHC(tot11,type = "HC1"))

tot12 = lm(etr ~ log1p(lob) + assets + log1p(emp) + log1p(roa), d)
summary(tot12)
coeftest(tot12,df=Inf,vcov=vcovHC(tot12,type = "HC1"))
```


```{r}
tot13 = lm(etr ~ lob + assets + emp + roa + rd, d_with_zero)
summary(tot13)
coeftest(tot13,df=Inf,vcov=vcovHC(tot13,type = "HC1"))

tot14 = lm(etr ~ lob + assets + emp + roa +rd, d_without_zero)
summary(tot14)
coeftest(tot14,df=Inf,vcov=vcovHC(tot14,type = "HC1"))

tot15 = lm(etr ~ log1p(lob) + assets + log1p(emp) + log1p(roa) + log1p(rd), d)
summary(tot15)
coeftest(tot15,df=Inf,vcov=vcovHC(tot15,type = "HC1"))
```
```{r}
tot16 = lm(etr ~ lob + assets + emp + roa + rd + levrat, d_with_zero)
summary(tot16)
coeftest(tot16,df=Inf,vcov=vcovHC(tot16,type = "HC1"))

tot17 = lm(etr ~ lob + assets + emp + roa +rd + levrat, d_without_zero)
summary(tot17)
coeftest(tot17,df=Inf,vcov=vcovHC(tot17,type = "HC1"))

tot18 = lm(etr ~ log1p(lob) + assets + log1p(emp) + log1p(roa) + log1p(rd) + log1p(levrat), d)
summary(tot18)
coeftest(tot18,df=Inf,vcov=vcovHC(tot18,type = "HC1"))
```

```{r}
tot19 = lm(etr ~ lob + assets + emp + roa + rd + levrat + caprat , d_with_zero)
summary(tot19)
coeftest(tot19,df=Inf,vcov=vcovHC(tot19,type = "HC1"))

tot20 = lm(etr ~ lob + assets + emp + roa +rd + levrat + caprat, d_without_zero)
summary(tot20)
coeftest(tot20,df=Inf,vcov=vcovHC(tot20,type = "HC1"))

tot21 = lm(etr ~ log1p(lob) + assets + log1p(emp) + log1p(roa) + log1p(rd) + log1p(levrat) +log1p(caprat), d)
summary(tot21)
coeftest(tot21,df=Inf,vcov=vcovHC(tot21,type = "HC1"))
```
```{r}
tot22 = lm(etr ~ lob + assets + emp + roa + rd + levrat + caprat + invrat , d_with_zero)
summary(tot22)
coeftest(tot22,df=Inf,vcov=vcovHC(tot22,type = "HC1"))

tot23 = lm(etr ~ lob + assets + emp + roa +rd + levrat + caprat +invrat, d_without_zero)
summary(tot23)
coeftest(tot23,df=Inf,vcov=vcovHC(tot23,type = "HC1"))

tot24 = lm(etr ~ log1p(lob) + assets + log1p(emp) + log1p(roa) + log1p(rd) + log1p(levrat) +log1p(caprat) +log1p(invrat), d)
summary(tot24)
coeftest(tot24,df=Inf,vcov=vcovHC(tot24,type = "HC1"))

```

```{r}
tot25 = lm(etr ~ lob + assets + emp + roa + rd + levrat + caprat + invrat , d_with_zero)
summary(tot25)
coeftest(tot25,df=Inf,vcov=vcovHC(tot25,type = "HC1"))

tot26 = lm(etr ~ log1p(lob) + assets + log1p(emp) + log1p(roa) + log1p(rd) + log1p(levrat) + log1p(caprat) + log1p(invrat), d)
summary(tot26)
coeftest(tot26,df=Inf,vcov=vcovHC(tot26,type = "HC1"))

```

Now table of the regressions: 

```{r}
stargazer(etrreg1, etrreg10, tot21, tot24, tot26, type = "text")
```


Running the same regressions but specifically on tax lobbying


```{r}
taxlob1 = lm(etr ~ taxlob, d_with_zero)
summary(taxlob1)
coeftest(taxlob1,df=Inf,vcov=vcovHC(taxlob1,type = "HC1"))

taxlob2 = lm(etr ~ taxlob, d_without_zero)
summary(taxlob2)
coeftest(taxlob2,df=Inf,vcov=vcovHC(taxlob2,type = "HC1"))

taxlob3 = lm(etr ~ log1p(taxlob), d)
summary(taxlob3)
coeftest(taxlob3,df=Inf,vcov=vcovHC(taxlob3,type = "HC1"))
```

```{r}
taxlob4 = lm(etr ~ taxlob + assets, d_with_zero)
summary(taxlob4)
coeftest(taxlob4,df=Inf,vcov=vcovHC(taxlob4,type = "HC1"))

taxlob5 = lm(etr ~ taxlob +assets , d_without_zero)
summary(taxlob5)
coeftest(taxlob5,df=Inf,vcov=vcovHC(taxlob5,type = "HC1"))

taxlob6 = lm(etr ~ log1p(taxlob) + assets, d)
summary(taxlob6)
coeftest(taxlob6,df=Inf,vcov=vcovHC(taxlob6,type = "HC1"))
```

```{r}
taxlob7 = lm(etr ~ taxlob + assets + emp, d_with_zero)
summary(taxlob7)
coeftest(taxlob7,df=Inf,vcov=vcovHC(taxlob7,type = "HC1"))

taxlob8 = lm(etr ~ taxlob + assets +emp, d_without_zero)
summary(taxlob8)
coeftest(taxlob8,df=Inf,vcov=vcovHC(taxlob8,type = "HC1"))

taxlob9 = lm(etr ~ log1p(taxlob) + assets + log1p(emp), d)
summary(taxlob9)
coeftest(taxlob9,df=Inf,vcov=vcovHC(taxlob9,type = "HC1"))
```
```{r}
taxlob10 = lm(etr ~ taxlob + assets + emp + rd, d_with_zero)
summary(taxlob10)
coeftest(taxlob10,df=Inf,vcov=vcovHC(taxlob10,type = "HC1"))

taxlob11 = lm(etr ~ taxlob + assets +emp +rd , d_without_zero)
summary(taxlob11)
coeftest(taxlob11,df=Inf,vcov=vcovHC(taxlob11,type = "HC1"))

taxlob12 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd), d)
summary(taxlob12)
coeftest(taxlob12,df=Inf,vcov=vcovHC(taxlob12,type = "HC1"))
```
```{r}
taxlob13 = lm(etr ~ taxlob + assets + emp + rd + levrat, d_with_zero)
summary(taxlob13)
coeftest(taxlob13,df=Inf,vcov=vcovHC(taxlob13,type = "HC1"))

taxlob14 = lm(etr ~ taxlob + assets +emp +rd + levrat , d_without_zero)
summary(taxlob14)
coeftest(taxlob14,df=Inf,vcov=vcovHC(taxlob14,type = "HC1"))

taxlob15 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(levrat), d)
summary(taxlob15)
coeftest(taxlob15,df=Inf,vcov=vcovHC(taxlob15,type = "HC1"))
```


```{r}
taxlob16 = lm(etr ~ taxlob + assets + emp + rd + roa + levrat, d_with_zero)
summary(taxlob16)
coeftest(taxlob16,df=Inf,vcov=vcovHC(taxlob16,type = "HC1"))

taxlob17 = lm(etr ~ taxlob + assets +emp +rd + roa + levrat , d_without_zero)
summary(taxlob17)
coeftest(taxlob17,df=Inf,vcov=vcovHC(taxlob17,type = "HC1"))

taxlob18 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat), d)
summary(taxlob18)
coeftest(taxlob18,df=Inf,vcov=vcovHC(taxlob18,type = "HC1"))
```
```{r}
taxlob19 = lm(etr ~ taxlob + assets + emp + rd + roa + levrat + caprat, d_with_zero)
summary(taxlob19)
coeftest(taxlob19,df=Inf,vcov=vcovHC(taxlob19,type = "HC1"))

taxlob20 = lm(etr ~ taxlob + assets +emp +rd + roa + levrat +caprat, d_without_zero)
summary(taxlob20)
coeftest(taxlob20,df=Inf,vcov=vcovHC(taxlob20,type = "HC1"))

taxlob21 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat), d)
summary(taxlob21)
coeftest(taxlob21,df=Inf,vcov=vcovHC(taxlob21,type = "HC1"))
```
```{r}
taxlob22 = lm(etr ~ taxlob + assets + emp + rd + roa + levrat + caprat + invrat, d_with_zero)
summary(taxlob22)
coeftest(taxlob22,df=Inf,vcov=vcovHC(taxlob22,type = "HC1"))

taxlob23 = lm(etr ~ taxlob + assets +emp +rd + roa + levrat +caprat +invrat, d_without_zero)
summary(taxlob23)
coeftest(taxlob23,df=Inf,vcov=vcovHC(taxlob23,type = "HC1"))

taxlob24 = lm(etr ~ log1p(taxlob) + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), d)
summary(taxlob24)
coeftest(taxlob24,df=Inf,vcov=vcovHC(taxlob24,type = "HC1"))
```
```{r}
taxlob25 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), d)
summary(taxlob25)
coeftest(taxlob25,df=Inf,vcov=vcovHC(taxlob25,type = "HC1"))
```

```{r}
stargazer(taxlob4, taxlob25,type = "text")
```

Adding fixed effects to the regressions 

```{r}
library("plm")
```


Running the same regressions before and after 2008 (world crisis)

```{r}
first = filter(d_with_zero, year <=2009) 
second = filter(d_with_zero, year >=2010)
```

Replace NA with 0's 

```{r}
d[is.na(d)] == 0
d

```

```{r}

```




Running regressions on original data set to compare how outliers impacted estimates for tax lobbying
```{r}
taxlob26 = lm(etr ~ taxlob + assets + emp + rd + roa + levrat + caprat + invrat, data)
summary(taxlob26)
coeftest(taxlob26,df=Inf,vcov=vcovHC(taxlob26,type = "HC1"))
```

```{r}
stargazer(taxlob25, taxlob26, type = "text")
```


Running regressions on original data set to compare how outliers impacted estimates for total lobbying

```{r}
tot27 = lm(etr ~ lob + assets + emp + roa + rd + levrat + caprat + invrat, data)
summary(tot27)
coeftest(tot27,df=Inf,vcov=vcovHC(tot27,type = "HC1"))
```

```{r}
stargazer(tot26, tot27, type = "text")
```

sector regressions
```{r}
sec=unique(d_without_zero$sector)
sec
sector1 = filter(d_with_zero, sector==sector[1])
sector2 = filter(d_with_zero, sector==sector[2])
sector3 = filter(d_with_zero, sector==sector[3])
sector4 = filter(d_with_zero, sector==sector[4])
sector5 = filter(d_with_zero, sector==sector[5])
sector6 = filter(d_with_zero, sector==sector[6])
sector7 = filter(d_with_zero, sector==sector[7])
sector8 = filter(d_with_zero, sector==sector[8])
sector9 = filter(d_with_zero, sector==sector[9])
sector10 = filter(d_with_zero, sector==sector[10])
sector11 = filter(d_with_zero, sector==sector[11])
sector12 = filter(d_with_zero, sector==sector[12])

sector13 = filter(d_without_zero, sector==sector[1])
sector14 = filter(d_without_zero, sector==sector[2])
sector15 = filter(d_without_zero, sector==sector[3])
sector16 = filter(d_without_zero, sector==sector[4])
sector17 = filter(d_without_zero, sector==sector[5])
sector18 = filter(d_without_zero, sector==sector[6])
sector19 = filter(d_without_zero, sector==sector[7])
sector20 = filter(d_without_zero, sector==sector[8])
sector21 = filter(d_without_zero, sector==sector[9])
sector22 = filter(d_without_zero, sector==sector[10])
sector23 = filter(d_without_zero, sector==sector[11])
sector24 = filter(d_without_zero, sector==sector[12])
```

Regressing Sectors:
Utility industry
```{r}
util = lm(etr ~ lob, sector1)
coeftest(util,df=Inf,vcov=vcovHC(util,type = "HC1"))
summary(util)
util2 = lm(etr ~  lob + assets + emp + rd + roa + levrat + caprat + invrat, sector1) 
summary(util2)
coeftest(util2,df=Inf,vcov=vcovHC(util2,type = "HC1"))

util3 = lm(etr ~log1p(lob) + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector1) 
summary(util3)
coeftest(util3,df=Inf,vcov=vcovHC(util3,type = "HC1"))

util4 = lm(etr ~ taxlob, sector1)
coeftest(util4,df=Inf,vcov=vcovHC(util4,type = "HC1"))
summary(util4)

util5 = lm(etr ~  taxlob + assets + emp + rd + roa + levrat + caprat + invrat, sector1) 
summary(util5)
coeftest(util5,df=Inf,vcov=vcovHC(util5,type = "HC1"))

util6 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector1) 
summary(util6)
coeftest(util6,df=Inf,vcov=vcovHC(util6,type = "HC1"))

```
IT industry
```{r}
it = lm(etr ~ lob, sector2)
coeftest(it,df=Inf,vcov=vcovHC(it,type = "HC1"))
summary(it)

it2 = lm(etr ~  lob + assets + emp + rd + roa + levrat + caprat + invrat, sector2) 
summary(it2)
coeftest(it2,df=Inf,vcov=vcovHC(util2,type = "HC1"))

it3 = lm(etr ~log1p(lob) + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector2) 
summary(it3)
coeftest(it3,df=Inf,vcov=vcovHC(it3,type = "HC1"))

it4 = lm(etr ~ taxlob, sector2)
coeftest(it4,df=Inf,vcov=vcovHC(it4,type = "HC1"))
summary(it4)

it5 = lm(etr ~  taxlob + assets + emp + rd + roa + levrat + caprat + invrat, sector2) 
summary(it5)
coeftest(it5,df=Inf,vcov=vcovHC(it5,type = "HC1"))

it6 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector2) 
summary(it6)
coeftest(it6,df=Inf,vcov=vcovHC(it6,type = "HC1"))
```

Health Care

```{r}
health = lm(etr ~ lob, sector3)
coeftest(health,df=Inf,vcov=vcovHC(health,type = "HC1"))
summary(health)

health2 = lm(etr ~ lob + assets + emp + rd + roa + levrat + caprat + invrat, sector3) 
summary(health2)
coeftest(health2,df=Inf,vcov=vcovHC(health2,type = "HC1"))

health3 = lm(etr ~log1p(lob) + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector3) 
summary(health3)
coeftest(health3,df=Inf,vcov=vcovHC(health3,type = "HC1"))

health4 = lm(etr ~ taxlob, sector3)
coeftest(health4,df=Inf,vcov=vcovHC(health4,type = "HC1"))
summary(health4)

health5 = lm(etr ~  taxlob + assets + emp + rd + roa + levrat + caprat + invrat, sector3) 
summary(health5)
coeftest(health5,df=Inf,vcov=vcovHC(health5,type = "HC1"))

health6 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector3) 
summary(health6)
coeftest(health6,df=Inf,vcov=vcovHC(health6,type = "HC1"))
```

Materials 

```{r}
material = lm(etr ~ lob, sector4)
coeftest(material,df=Inf,vcov=vcovHC(util,type = "HC1"))
summary(material)

material2 = lm(etr ~  lob + assets + emp + rd + roa + levrat + caprat + invrat, sector4) 
summary(material2)
coeftest(material2,df=Inf,vcov=vcovHC(material2,type = "HC1"))

material3 = lm(etr ~log1p(lob) + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector4) 
summary(material3)
coeftest(material3,df=Inf,vcov=vcovHC(material3,type = "HC1"))

material4 = lm(etr ~ taxlob, sector4)
coeftest(material4,df=Inf,vcov=vcovHC(util4,type = "HC1"))
summary(material4)

material5 = lm(etr ~  taxlob + assets + emp + rd + roa + levrat + caprat + invrat, sector4) 
summary(material5)
coeftest(material5,df=Inf,vcov=vcovHC(material5,type = "HC1"))

material6 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector4) 
summary(material6)
coeftest(material6,df=Inf,vcov=vcovHC(material6,type = "HC1"))

```

Energy 
```{r}
energy = lm(etr ~ lob, sector5)
coeftest(energy,df=Inf,vcov=vcovHC(energy,type = "HC1"))
summary(energy)

energy2 = lm(etr ~  lob + assets + emp + rd + roa + levrat + caprat + invrat, sector5) 
summary(energy2)
coeftest(energy2,df=Inf,vcov=vcovHC(energy2,type = "HC1"))

energy3 = lm(etr ~log1p(lob) + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector5) 
summary(energy3)
coeftest(energy3,df=Inf,vcov=vcovHC(energy3,type = "HC1"))

energy4 = lm(etr ~ taxlob, sector5)
coeftest(energy4,df=Inf,vcov=vcovHC(energy4,type = "HC1"))
summary(energy4)

energy5 = lm(etr ~  taxlob + assets + emp + rd + roa + levrat + caprat + invrat, sector5) 
summary(energy5)
coeftest(energy5,df=Inf,vcov=vcovHC(energy5,type = "HC1"))

energy6 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector5) 
summary(energy6)
coeftest(energy6,df=Inf,vcov=vcovHC(energy6,type = "HC1"))
```

Transport
```{r}
tran = lm(etr ~ lob, sector6)
coeftest(tran,df=Inf,vcov=vcovHC(tran,type = "HC1"))
summary(tran)

tran2 = lm(etr ~  lob + assets + emp + rd + roa + levrat + caprat + invrat, sector6) 
summary(tran2)
coeftest(tran2,df=Inf,vcov=vcovHC(tran2,type = "HC1"))

tran3 = lm(etr ~log1p(lob) + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector6) 
summary(tran3)
coeftest(tran3,df=Inf,vcov=vcovHC(tran3,type = "HC1"))

tran4 = lm(etr ~ taxlob, sector6)
coeftest(tran4,df=Inf,vcov=vcovHC(tran4,type = "HC1"))
summary(tran4)

tran5 = lm(etr ~  taxlob + assets + emp + rd + roa + levrat + caprat + invrat, sector6) 
summary(tran5)
coeftest(tran5,df=Inf,vcov=vcovHC(tran5,type = "HC1"))

tran6 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector6) 
summary(tran6)
coeftest(tran6,df=Inf,vcov=vcovHC(tran6,type = "HC1"))
```


Industrials

```{r}
indus = lm(etr ~ lob, sector7)
coeftest(indus,df=Inf,vcov=vcovHC(indus,type = "HC1"))
summary(indus)

indus2 = lm(etr ~  lob + assets + emp + rd + roa + levrat + caprat + invrat, sector7) 
summary(indus2)
coeftest(indus2,df=Inf,vcov=vcovHC(indus2,type = "HC1"))

indus3 = lm(etr ~log1p(lob) + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector7) 
summary(indus3)
coeftest(indus3,df=Inf,vcov=vcovHC(indus3,type = "HC1"))

indus4 = lm(etr ~ taxlob, sector7)
coeftest(indus4,df=Inf,vcov=vcovHC(indus4,type = "HC1"))
summary(indus4)

indus5 = lm(etr ~  taxlob + assets + emp + rd + roa + levrat + caprat + invrat, sector7) 
summary(indus5)
coeftest(indus5,df=Inf,vcov=vcovHC(indus5,type = "HC1"))

indus6 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector7) 
summary(indus6)
coeftest(indus6,df=Inf,vcov=vcovHC(indus6,type = "HC1"))
```

Finance Regressions

```{r}
fin = lm(etr ~ lob, sector8)
coeftest(fin,df=Inf,vcov=vcovHC(fin,type = "HC1"))
summary(fin)

fin2 = lm(etr ~  lob + assets + emp + rd + roa + levrat + caprat + invrat, sector8) 
summary(fin2)
coeftest(fin2,df=Inf,vcov=vcovHC(fin2,type = "HC1"))

fin3 = lm(etr ~log1p(lob) + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector8) 
summary(fin3)
coeftest(fin3,df=Inf,vcov=vcovHC(fin3,type = "HC1"))

fin4 = lm(etr ~ taxlob, sector8)
coeftest(fin4,df=Inf,vcov=vcovHC(fin4,type = "HC1"))
summary(fin4)

fin5 = lm(etr ~  taxlob + assets + emp + rd + roa + levrat + caprat + invrat, sector8) 
summary(fin5)
coeftest(fin5,df=Inf,vcov=vcovHC(fin5,type = "HC1"))

fin6 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector8) 
summary(fin6)
coeftest(fin6,df=Inf,vcov=vcovHC(fin6,type = "HC1"))
```

Consumer Discretionary 

```{r}
cd = lm(etr ~ lob, sector9)
coeftest(cd,df=Inf,vcov=vcovHC(cd,type = "HC1"))
summary(cd)

cd2 = lm(etr ~  lob + assets + emp + rd + roa + levrat + caprat + invrat, sector9) 
summary(cd2)
coeftest(cd2,df=Inf,vcov=vcovHC(cd2,type = "HC1"))

cd3 = lm(etr ~log1p(lob) + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector9) 
summary(cd3)
coeftest(cd3,df=Inf,vcov=vcovHC(cd3,type = "HC1"))

cd4 = lm(etr ~ taxlob, sector9)
coeftest(cd4,df=Inf,vcov=vcovHC(cd4,type = "HC1"))
summary(cd4)

cd5 = lm(etr ~  taxlob + assets + emp + rd + roa + levrat + caprat + invrat, sector9) 
summary(cd5)
coeftest(cd5,df=Inf,vcov=vcovHC(cd5,type = "HC1"))

cd6 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector9) 
summary(cd6)
coeftest(cd6,df=Inf,vcov=vcovHC(cd6,type = "HC1"))
```

Real Estate Regressions 

```{r}
re = lm(etr ~ lob, sector10)
coeftest(re,df=Inf,vcov=vcovHC(cd,type = "HC1"))
summary(re)

re2 = lm(etr ~  lob + assets + emp + rd + roa + levrat + caprat + invrat, sector10) 
summary(re2)
coeftest(re2,df=Inf,vcov=vcovHC(re2,type = "HC1"))

re3 = lm(etr ~log1p(lob) + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector10) 
summary(re3)
coeftest(re3,df=Inf,vcov=vcovHC(re3,type = "HC1"))

re4 = lm(etr ~ taxlob, sector10)
coeftest(re4,df=Inf,vcov=vcovHC(re4,type = "HC1"))
summary(re4)

re5 = lm(etr ~  taxlob + assets + emp + rd + roa + levrat + caprat + invrat, sector10) 
summary(re5)
coeftest(re5,df=Inf,vcov=vcovHC(re5,type = "HC1"))

re6 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector10) 
summary(re6)
coeftest(re6,df=Inf,vcov=vcovHC(re6,type = "HC1"))
```

Consumer Staples Regressions 

```{r}
cs = lm(etr ~ lob, sector11)
coeftest(cs,df=Inf,vcov=vcovHC(cs,type = "HC1"))
summary(cs)

cs2 = lm(etr ~  lob + assets + emp + rd + roa + levrat + caprat + invrat, sector11) 
summary(cs2)
coeftest(cs2,df=Inf,vcov=vcovHC(cs2,type = "HC1"))

cs3 = lm(etr ~log1p(lob) + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector11) 
summary(cs3)
coeftest(cs3,df=Inf,vcov=vcovHC(cs3,type = "HC1"))

cs4 = lm(etr ~ taxlob, sector11)
coeftest(cs4,df=Inf,vcov=vcovHC(cs4,type = "HC1"))
summary(cs4)

cs5 = lm(etr ~  taxlob + assets + emp + rd + roa + levrat + caprat + invrat, sector11) 
summary(cs5)
coeftest(cs5,df=Inf,vcov=vcovHC(cs5,type = "HC1"))

cs6 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector11) 
summary(cs6)
coeftest(cs6,df=Inf,vcov=vcovHC(cs6,type = "HC1"))

```

Communications Regressions

```{r}
commun = lm(etr ~ lob, sector12)
coeftest(commun,df=Inf,vcov=vcovHC(commun,type = "HC1"))
summary(commun)

commun2 = lm(etr ~  lob + assets + emp + rd + roa + levrat + caprat + invrat, sector12) 
summary(commun2)
coeftest(commun2,df=Inf,vcov=vcovHC(commun2,type = "HC1"))

commun3 = lm(etr ~log1p(lob) + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector12) 
summary(commun3)
coeftest(commun3,df=Inf,vcov=vcovHC(commun3,type = "HC1"))

commun4 = lm(etr ~ taxlob, sector12)
coeftest(commun4,df=Inf,vcov=vcovHC(commun4,type = "HC1"))
summary(commun4)

commun5 = lm(etr ~  taxlob + assets + emp + rd + roa + levrat + caprat + invrat, sector12) 
summary(commun5)
coeftest(commun5,df=Inf,vcov=vcovHC(commun5,type = "HC1"))

commun6 = lm(etr ~ taxlob + assets + log1p(emp) + log1p(rd) + log1p(roa) + log1p(levrat) + log1p(caprat) + log1p(invrat), sector12) 
summary(commun6)
coeftest(commun6,df=Inf,vcov=vcovHC(commun6,type = "HC1"))

```


Dummy variables 
```{r}
 data$dummyYear <- as.numeric(data$year == 2000& 2004 &2008 & 2012 & 2016)
```

